---
title: "(Meta)data texting in `{ggplot2}`"
subtitle: "Live-code"
author: "Corrado Lanera"
date: "lesson: 2020-09-14 (last update: `r Sys.Date()`)"
output:
  html_notebook:
      toc: true
      toc_float: true
---

# Setup

```{r setup, message = FALSE}
library(tidyverse)
library(palmerpenguins)
```


# Data

We are interested to use `penguins` data for our plot:

```{r}
citation("palmerpenguins")
penguins
```










# Aesthetics

We would like to map at `bill_length_mm` (in `x`) vs `bill_depth_mm` (in
`y`) the `sex` of the penguins, coloured by their `species`. 

```{r}
penguins
```










# Geoms

## Classic

We would like to see in the plot the text reporting pengiuns' sex like
in a scatterplot. 

Considering to adjust transparency and size for a better visualization
for overlapping labels.

```{r}
penguins
```





## Annotations

We would also put two tags inside the plot as a reference for zone of
stumpy and tapered penguins in the top-left and bottom-right corners 
respectively.

```{r}
penguins
```










# Legends (Scales)

## Legends

We need to write in better way bot the title of the axis and the text 
of the axes' ticks (possibly reporting the unit of measure).

Also, write in al UPPERCASE the spieces in the legend, and try to 
use colors coherently with the penguin's artwork by
[@allison_horst](https://www.allisonhorst.com/).

You can use the following colours' codes:

 - `#C55ACA`: Chinstrap
 - `#0E7274`: Gentoo
 - `#FF6F04`: Adelie
 
```{r}
penguins
```





## `labs()`

Now we can put a title ("penguins' bills across islands"), a caption
("Data from {palmerpenguins} package."), and fix the title of the
colour's legend (capitalize the first letter).

```{r}
penguins
```










# Facets

We would like to draw this plot for each island in the same graph, in
a grid of longwise plots (ie, three rows). We will love if in the
strips' header will compare the indication that what is reporte is 
actually the name of an island.

```{r}
penguins
```










# Theme

Finally we twant to take care of the style of our plot:

 - Set a bigger fon size for the title (eg 24px)
 - Set the species names reported in the legend in monospaced characters
 - Set the ticks' labels for both the axes in blue
 - Set the strips' headers in boldface
 
```{r}
penguins
```
 
 
 
