---
title: "(Meta)data texting in `{ggplot2}`"
subtitle: "Check `{ggplot2}` basics"
author: "Corrado Lanera"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: false
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(parsons)
library(tidyverse)
library(palmerpenguins)

knitr::opts_chunk$set(echo = FALSE)
```


## Checks

This introductory set of quiz aims to check the basic `{ggplot}`
requirements for the class. In case of doubt, feel free to explore also
the wrong answers: short feedback useful to understand why the answer is
wrong will be displayed.

### Data

> Info: Data are _the_ information you want to display.

Answer the following question.

```{r base-data}
quiz(caption = "Data-quiz",
  question("What can you use as data input to `ggplot()`",
           
    answer("A `data.frame` (maybe a _tibble_)",
           correct = TRUE,
           message = random_praise()),
    
    answer("One vector for each dimension to plot",
           message = "`ggplot()` require a single object."),
    
    answer("A `matrix`",
           message = "What if you need columns of distinct types?"),
    
    answer("A tibble only (standard `data.frames` must be converted)",
           message = "The Tidyverse is an inclusive place ;-)"),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)

```

### Aesthetics

> Info: Aesthetics are maps from the (variables into the) data to the 
corresponding parts of the plot they will be used for.

Answer the following question.

```{r base-aes}
quiz(caption = "Aesthetics-quiz",
  question("What are the first two default aesthetics/input of `aes()`",
           
    answer("`x` and `y` (both optional)",
           correct = TRUE,
           message = random_praise()),
    
    answer("There is no default, you must name all the aesthetics you require (`x` and `y` included)",
           message = "`x` and `y` are really sooo common ;-)."),
    
    answer("`x` and `y` (at least one mandatory)",
           message = "What if you would like to inherits previous `x` and `y` aesthetics (e.g. into a `geom_*()`), and set only some new ones?"),
    
    answer("`x` and `y` (both mandatory)",
           message = "How many aesthetics require `geom_density()` to plot a distribution?"),
    

    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
    

  question("What is the expected input for `x` in `aes(x = ___)`?",
           
    answer("The unquoted name of a variable from the data used (eg, `my_var`)",
           correct = TRUE,
           message = random_praise()),
    
    answer("The quoted name of a variable from the data used (eg, `\"my_var\"`",
           message = "This won't throw an error, but generally won't produce the output you expect (_de facto_ `\"my_var\"` is a single, textual, data)"),
    
    answer("A vector of values to map into the x-axis",
           message = "You _can_ use a vector to map explicitely into an aesthetic, but you don't do that inside `aes()`"),
    

    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```


### Geoms

> Info: Geoms are the graphical objects used in the plot to draw the information
in your data in the way described by the aesthetics.

Answer the following question. Next, drag and drop a valid set of chunks
from the right box to the left one to define a valid syntax for a
`{ggplot2}` plot. Note: not all the purposed component are valid, but
there are two valid solutions (It is enough that you find one of them).

```{r base-geom}
quiz(caption = "Geom-quiz",
  question("Can you use more than one `geom_*()` drawing a plot?",
           
    answer("Yes, and even with multiple instances of the same `geom_*()`",
           correct = TRUE,
           message = random_praise()),
    
    answer("No, you cannot: a plot can reports only a single geometry",
           message = "Think about a scatter plot with their regression line superimposed... how many geometries are there?"),
    
    answer("Yes, but only if all the `geom_*()`s are distinct each-other",
           message = "Why you don't want to be able to disply, eg,  multiple regression line using distinct methods for the same set of points?"),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  
  question_parsons(
    initial = c(
      "ggplot(penguins)",
      "ggplot(penguins, aes(year, body_mass_g))",
      "ggplot(penguins, aes(\"year\", \"body_mass_g\"))",
      "ggplot(penguins, \"year\", \"body_mass_g\")",
      "ggplot(penguins, year, body_mass_g)",
      "ggplot(penguins$year, penguins$body_mass_g)",
      "geom_point()",
      "geom_point(year, body_mass_g)",
      "geom_point(aes(year, body_mass_g))"
    ),
    answer(
      c(
        "ggplot(penguins, aes(year, body_mass_g))",
        "geom_point()"
      ),
      correct = TRUE,
      message = random_praise()
    ),
    answer(
      c(
        "ggplot(penguins)",
        "geom_point(aes(x = year, y = body_mass_g))"
      ),
      correct = TRUE,
      message = random_praise()
    ),

    orientation = "vertical",
    problem_type = "ggplot2",
    try_again = random_encouragement(),
    submit_button = "Submit Answer",
    allow_retry = TRUE
  )
)
```


### Well done! {data-allow-skip=TRUE}

This quick check is finished, and it was a great success!
Congratulation! You can be confident in following the introductory 
class on using text within `ggplot()`s!

Now, feel free to explore your own code and solutions here playing with 
penguins and `{ggplot2}` as much as you like!

```{r base-explore, exercise=TRUE, exercise.lines=5}
library(tidyverse)
library(palmerpenguins)



```












```{r quiz-template, eval=FALSE, include=FALSE}
quiz(caption = "Geom-quiz",
  question("",
           
    answer("",
           correct = TRUE,
           message = random_praise()),
    
    answer("",
           message = ""),
    
    answer("",
           message = ""),
    
    answer("",
           message = ""),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```



```{r parsons-template, eval=FALSE, include=FALSE}
question_parsons(
  initial = c(
    "iris",
    "mutate(...)",
    "summarize(...)",
    "print()"
  ),
  pass_if(
    c(
      "iris",
      "mutate(...)",
      "summarize(...)"
    )
  ),
  fail_if(
    ~length(.) < 2,
    message = "Include at least two answers"
  ),
  fail_if(
    function(x){"print()" %in% x},
    message = "You should not include print() in your answer"
  ),
  fail_if(
    ~{.[1] != "iris"},
    message = "Your solution should start with 'iris'"
  )
)
```


