---
title: "(Meta)data texting in `{ggplot2}`"
subtitle: "Formative assessments"
author: "Corrado Lanera"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: false
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(vdiffr)
library(tidyverse)
library(palmerpenguins)

knitr::opts_chunk$set(echo = FALSE)
gradethis::gradethis_setup()


# https://github.com/rstudio-education/gradethis/issues/97
test_plots <- function(x, y) {
  fx <- fs::file_temp()
  fy <- fs::file_temp()
  # leave no footprint
  on.exit({
    unlink(c(fx, fy))
  })
  # write x
  vdiffr::write_svg(x, fx)
  # write y
  vdiffr::write_svg(y, fy)
  # compare
  identical(readLines(fx), readLines(fy))
}

```


## Step-by-step checks

### Data

Answer the following questions.

```{r quiz-data}
quiz(caption = "Data-quiz",
  question("Textual information you found for each observation into your
           data are:",
           
    answer("data",
           correct = TRUE,
           message = random_praise()),
    
    answer("meta-data",
           message = "Meta-data are a properties of the data you have,
           not the data itself"),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  
  question("When you want to add a new (i.e., not already contained in
           your main data) single piece of text inside a plot, it is:",
           
    answer("data",
           correct = TRUE,
           message = paste(random_praise(), "It is a single new data
             point. You can create a new data frame with a single row 
             and pass this data to a suitable geoms, and continue to
             work in the standard primary way. On the other hand, the 
             `annotation()` layer is provided for precisely this reason,
             and it is much much quicker and simpler to use!")),
    
    answer("meta-data",
           message = "Meta-data are properties of the data you have,
           not the data itself. In this case this piece of data are not
           a data included in your main dataset, but it still a
           (single-instance of new) data."),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```



### Aesthetics

Answer the following question.

```{r quiz-aes}
quiz(caption = "Aes-quiz",
  question("Suppose you have a variable `var` reporting textual 
           information for each observation in the data. You would like
           to plot those pieces of information into the plot at the
           corresponding coordinates indicated by the variables `x`,
           and `y` (already in your data).
           How would you set the main aesthetics into the `ggplot()`
           call?",
           
    answer("`aes(x, y, label = var)`",
           correct = TRUE,
           message = random_praise()),
    
    answer("`aes(x, y, text = var)`",
           message = "geoms drawing text does not use the keyword
           `text` for that but `label`"),
    
    answer("`aes(x, y, var)`",
           message = "Only the first two aesthetics (`x`, and `y`) are
           positional, ie, you won't require to set them by name."),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

### {data-allow-skip=TRUE}
Check for yourself

```{r quiz-aes-code, exercise = TRUE, eval=FALSE}
penguins %>% 
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, ___ = species)) +
  geom_text()
```





### Geom

Answer the following questions (in the first one you can select more
than one correct answer).

```{r quiz-geom}
quiz(caption = "Geom-quiz",
  question("Which `geom_*`s are suitable for text?",
           
    answer("`geom_text()`",
           correct = TRUE,
           message = random_praise()),
    
    answer("`geom_label()`",
           correct = TRUE,
           message = paste(random_praise(), "`geom_label()` is like
           `geom_text()` but it draws a rectangle behind the text,
           making it easier to read")),
    
    answer("`geom_point()`",
           message = "The text will be used as points in a scatterplot, 
           but they are not \"point\", they are text, sometimes called
           labels. If you call `geom_point()` you will obtain a simple
           point at every set of coordinates (x, y) in the data, without 
           text or labels."),
    
    answer("`geom_word()`",
           message = "This geom does not exist in `{ggplot2}`
           currently."),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  
  
  question("If you want to draw ALL your textual data in blue, How can
           you set the call to `geom_text()`?",
           
    answer("`geom_text(colour = \"blue\")`",
           correct = TRUE,
           message = random_praise()),
    
    answer("`geom_text(aes(colour = \"blue\"))`",
           message = "Surprisingly, this will create a new aesthetic map 
           to draw the **data** \"blue\" (which is constant) to the 
           colour property. So this is not what you want, ie, **set** a
           custom property without any data map. Surprisingly, the call 
           will draw all your data in red :-)... if you are curious 
           about the why you can take a look at chapter 12.4 of the 
           [ggplot2 book](https://ggplot2-book.org/)"),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

### {data-allow-skip=TRUE}
Check for yourself

```{r quiz-geom-code, exercise = TRUE, eval=FALSE}
penguins %>% 
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, label = species)) +
  geom____() # try: colour = "blue" / aes(colour = "blue") ?
```


### Annotation

Answer the following questions (in the first one you can select more
than one correct answer).

```{r quiz-annotation}
quiz(caption = "Geom-quiz",
  question("What is the first argument of `annotate()`?",
           
    answer("A character string matching a (any) geom, like \"text\"",
           correct = TRUE,
           message = paste(random_praise(), "Yes, `annotate()` can draw
             any explicitly provided data point(s) using any geoms.
             Particularly useful but not limited to text and labels")),
    
    answer("A character string of the text annotation to plot",
           message = "`annotate()` is a wrapper for a single occurrence 
           of a geom. You need to first decide the geom to know which
           aesthetics to use. Next, you will explicitly set those 
           aesthetics without any data mapping."),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

### {data-allow-skip=TRUE}
Check for yourself

```{r quiz-annotation-code, exercise = TRUE, eval=FALSE}
penguins %>% 
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, label = species)) +
  geom_text(colour = "blue") +
  annotate(___, x = 45, y = 17.5, label = "center", # try both "text", and "label"
           colour = "red", fill = "blue", size = 24)
```



  





### Scales

Answer the following question (select all the answers you think are
correct)

```{r quiz-scales}
quiz(caption = "Scale-quiz",
  question("What labeling functions you can use from the package
           `{scale}` (automatically imported with `{ggplot2}`) to 
           format ticks' labels for the axes? (Note: you are not supposed to know the correct answer(s) to this question...just guess it/them ;-))",
           
    answer("`label_number()` for decimal format (e.g. 0.12, 1,234)",
           correct = TRUE,
           message = random_praise()),
    answer("`label_scientific()` for scientific notation (e.g. 1e05, 1.5e-02)",
           correct = TRUE,
           message = random_praise()),
    answer("`label_bytes()` to scale bytes into human friendly units (e.g. 1 kb, 2 MB, etc)",
           correct = TRUE,
           message = random_praise()),
    answer("`label_dollar()` to label currencies (i.e. not only dollars)",
           correct = TRUE,
           message = random_praise()),
    answer("`label_percent()` for % to the number (e.g. 2.5%, 50%, etc)",
           correct = TRUE,
           message = random_praise()),
    answer("`label_pvalue()` use \"<\" and \">\" for p-values close to 0 and 1 (e.g. <0.001, 0.25, p >= 0.99)",
           correct = TRUE,
           message = random_praise()),
    answer("`label_number_si()` use SI prefixes (e.g. 2k, 1M, 5T etc)",
           correct = TRUE,
           message = random_praise()),
    answer("`label_wrong()` for answer suggestion (e.g. \"check this too: wrong!`\")",
           message = "Can we suppose that `label_wrong()` could be...wrong? :-)"),

        
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  
  
  question("How can you set the title of the colour legend to \"New
           title\"?",
           
    answer("Using `labs(colour = \"New title\")`",
           correct = TRUE,
           message = random_praise()),
    
    answer("Using `labs(legend = \"New title\")`",
           message = "You can have more than one legend in a plot,
           so using `labs(legend = \"New title\")` would represent an 
           undefidend intent"),
    
    answer("`scale_colour_<TYPE>(\"New title\")`",
           correct = TRUE,
           message = random_praise()),
    
    answer("`ggplot(aes(x, y, colour = \"New title\"))`",
           message = "This will be change the whole colour aesthetics
           to use a single colour (red, but named \"New title\") with
           a corresponding legend's title named \"colour\". You need to
           change the name of the legend only, and you have two options:
           using the first argument of the corresponding
           `scale_<AES>_<TYPE>()` call, or to call `labs()` (which aims
           to set all the legends'names) and pass your title to its 
           `colour` argument"),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

### {data-allow-skip=TRUE}
Check for yourself

```{r quiz-scales-code, exercise = TRUE, eval=FALSE}
penguins %>% 
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm,
             label = species, colour = island)) +
  geom_point() +
  scale_x_continuous(labels = scales::label____()) +
  scale_y_continuous(labels = scales::label____()) +
  labs(___ = ___)
```









### Facets

Answer the following question.

```{r quiz-facet}
quiz(caption = "Facet-quiz",
  question("How can you personalize the titles of the facets in a grid
           plot?",
           
    answer("Using the `labeller` option of `facet_*()`",
           correct = TRUE,
           message = random_praise()),
    
    answer("Changing the content of the facetting varible(s) in the
           input data frame.",
           message = "This indeed works, but use a `labeller` would be 
           much better. Take a look at `?labeller` and  
           http://bit.ly/change-facets-titles"),
    
    answer("Using `labs(facet = ___)`",
           message = "`labs()` set only titles! The names of the 
           facet's strips could seems like titles but they are actually 
           the \"text\" of that special kind of legends (ie, they are
           the levels of the content of the faceting
           variables, not their names!). On the other hand, keep in mind
           that facets are not managed by scales function as it does 
           with all the other legends, but they require a special 
           labeller called directly inside the `facet_*()` call."),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

### {data-allow-skip=TRUE}
Check for yourself (you have some `hints` here)

```{r quiz-facet-code, exercise = TRUE, eval=FALSE}
penguins %>% 
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, label = species)) +
  geom_point() +
  facet_grid(island~sex, labeller = ___)
```


```{r quiz-facet-code-hint, eval=FALSE}
# try
facet_grid(island~sex, labeller = "label_both")

# or try
facet_grid(island~sex,
           labeller = labeller(sex = c(female = "F", male = "M")))

# or try
facet_grid(island~sex,
           labeller = labeller(island = label_both, sex = label_value))
```









### Theme

Answer the following questions (you can select more than one correct
answer in the first and the second-to-last ones).

```{r quiz-theme}
quiz(caption = "Theme-quiz",
  question("What can you change using `theme()`?", type = "multiple", 
           
    answer("Style",
           correct = TRUE,
           message = random_praise()),
    
    answer("Data",
           message = "`theme()` involves style only, no data change.
           Using `theme()` you can only change _how_ things are viewed,
           not _what_!"),
    
    answer("Meta-data",
           message = "`theme()` involves style only, no meta-data change.
           Using `theme()` you can only change _how_ things are viewed,
           not _what_!"),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  
  question("What would you use to change the style of something
           related to the whole plot (like the font family of its
           title)",
           
    answer("A call to `plot.*` inside `theme()`",
           correct = TRUE,
           message = random_praise()),
    
    answer("Setting, eg, the `title.family` option into the `labs()` call",
           message = "Using `labs()` you can set the title (with `title = `),
           but it is used to define it only, not its style."),
    answer("Setting the aesthetic `family` in the main `aes()` into the
           call to `ggplot()`",
           message = "An aesthetic maps information inside the data, do 
           not set how a piece of mapped information would be styled"),

    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  
  question("What would you use to change the style of something
           related to the faceting's strips?",
           
    answer("A call to `strip.*` inside `theme()`",
           correct = TRUE,
           message = random_praise()),
    
    answer("Setting a `family` property into a customized `labeller` 
           function to use inside the `facet_*()` call",
           message = "A `labeller` produce _what_ will be displayed 
           into the faceting strips' headers, i.e. it change the data
           to use for them accordingly to your interest. Anyway, it does
           not involve style, only the data."),
    answer("A call to `facet.*` inside `theme()`", message = "Unfortunately, `facet.*` is not how 
           `{ggplot2}` call facets when referring to them. But, they
           look like strips..."),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  
  question("What can you use to change the style of both the axes' names 
           in the same way??",
           
    answer("`axis.title`",
           correct = TRUE,
           message = random_praise()),
    
    answer("set `axis.title.x`, `axis.title.y` in the same way",
           correct = TRUE,
           message = paste(random_praise(), "Anyway, \"set
           `axis.title.x`, `axis.title.y` in the same way\" would not be 
             the best choice if the changes are the same.")),
    answer("`axis.text`"),
    answer("set `axis.text.x`, `axis.text.y` in the same way"),

    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  
  question("What can you set using `element_text()`?",
           
    answer("Only the style of the corresponding text: font face, family,
           size, angle, colour, etc",
           correct = TRUE,
           message = random_praise()),
    
    answer("Only textual elements: like lables, and their position",
           message = paste(random_praise(), "`theme()` involves style
           only, no data nor meta-data change. Using `theme()` you can
           only change _how_ things are viewed, not _what_!")),
    answer("All the element involved in plotting text: fromt the
           elements displayed to their style",
           message = paste(random_praise(), "`theme()` involves style
           only, no data nor meta-data change. Using `theme()` you can
           only change _how_ things are viewed, not _what_!")),

    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```


### {data-allow-skip=TRUE}
Check for yourself (you have some `hints` here)

```{r quiz-theme-code, exercise = TRUE, eval=FALSE}
penguins %>% 
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm,
             label = species, colour = island)) +
  geom_text() +
  labels(___ = "Title")
  theme(plot.___ = ___(___ = "bold"),
        legend.title = ___(___ = "darkgreen"),
        axis.___.x = ___(angle = -30))
```


### {data-allow-skip=TRUE}
Wonderful!! You have completed all the step-by-step assessment for this
class. You are now ready to take the _Final Test_.

Feel free to play with penguins and `{ggplot2}` as much as you like! 
When you are ready, press `Continue`/`Next Topic`.

```{r thopic-free, exercise=TRUE, exercise.lines = 5}
library(tidyverse)
library(palmerpenguins)



```



## Final Test



###

In the following, you will be asked to create a plot similar to the one
created in the class. We will reach the final result step by step!

In the following, you will be required to modify/update a purposed code
(similar to the one used in the class) accordingly to some instruction.

You can experiment with your code as much as you want to click on the
`Run Code` button. You can click on `Start Over` to reset the code chunk
to the original purposed one. You can click to `Hints` to come across
some hints if you need them.

When you think you have reached the correct result, you can click on the
`Submit Answer` to check it.

> NOTE: if your answer is wrong, you can try again as much time as you
like. But you must pass a stage to go to the next one (there is only one
stage marked as "optional", which you can skip if you won't pass it).


Ready!? Let's go!



### Data and aesthetics

Map `body_mass_g` against `flipper_length_mm` showing the `species` at 
each data-point, coloured by `sex`.

```{r final-data, exercise=TRUE, exercise.lines = 4}
penguins %>% 
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm,
    label = sex, colour = species)) +
  geom_text(check_overlap = TRUE)
```

```{r final-data-check}
grade_result(pass_if(~{
  ref <- penguins %>% 
    ggplot(aes(x = body_mass_g, y = flipper_length_mm,
      label = species, colour = sex)) +
    geom_text(check_overlap = TRUE)

  test_plots(.result, ref)
  
}), incorrect = "Wrong answer.")
```


```{r final-data-hint-1, eval = FALSE}
# modify only (you do not need to add or remove lines of code)
aes(x = ___, y = ___, label = ___, colour = ___)
```

```{r final-data-hint-2, eval = FALSE}
# modify only (you do not need to add or remove lines of code)
aes(x = body_mass_g, y = flipper_length_mm,
    label = species, colour = sex)
```



### Annotations (_una tantum_ data)

Draw a label in the middle-left and in the middle-right part of the plot
reporting "Short" and "Long" respectively (use the "inward" options as
in class for a nicer (and graded correctly) result)

```{r final-geom, exercise=TRUE, exercise.lines = 8}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = flipper_length_mm,
    label = species, colour = sex)) +
  geom_text(check_overlap = TRUE)
```


```{r final-geom-check}
grade_result(pass_if(~{
  ref <- penguins %>% 
    ggplot(aes(x = body_mass_g, y = flipper_length_mm,
      label = species, colour = sex)) +
    geom_text(check_overlap = TRUE) +
    annotate("label", x = 4500, y = 170, label = "Short",
             hjust = "inward", vjust = "inward") +
    annotate("label", x = 4500, y = 230, label = "Long",
             vjust = "inward", hjust = "inward")


  test_plots(.result, ref)
  
}), incorrect = "Wrong answer.")
```

```{r final-geom-hint-1, eval = FALSE}
# use the following coordinate for the annotations
x = 4500, y = 170 # short
x = 4500, y = 230 # long
```

```{r final-geom-hint-2, eval = FALSE}
# use the following coordinate for the annotations
x = 4500, y = 170 # short
x = 4500, y = 230 # long

# the template for (one) annotation is the following one
annotate("label", x = ___, y = ___, label = ___,
         hjust = "inward", vjust = "inward")
```

```{r final-geom-hint-3, eval = FALSE}
# use the following coordinate for the annotations
x = 4500, y = 170 # short
x = 4500, y = 230 # long

# the template for (two) annotations is the following one
annotate("label", x = ___, y = ___, label = ___,
         hjust = "inward", vjust = "inward") +
annotate("label", x = ___, y = ___, label = ___,
         hjust = "inward", vjust = "inward")

# Remember to capitalize short ("Short") and long ("Long") for your
# label.
```


 

> Optional: Just for fun, try to remove all the `invard`s options from
your final (correct) result.


```{r final-geom-2, exercise=TRUE, exercise.lines = 6}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = flipper_length_mm,
    label = species, colour = sex)) +
  geom_text(check_overlap = TRUE)
```


```{r final-geom-2-check}
grade_result(
  pass_if(~{
    ref <- penguins %>% 
      ggplot(aes(x = body_mass_g, y = flipper_length_mm,
        label = species, colour = sex)) +
      geom_text(check_overlap = TRUE) +
      annotate("label", x = 4500, y = 170, label = "Short") +
      annotate("label", x = 4500, y = 230, label = "Long")
  
      test_plots(.result, ref)
  
  }),
  fail_if(~TRUE, paste(
    random_encouragement(),
    "(Anyway, you can go on: this is not required to be correct!)"))
)
```





### Scales - axes (Meta-data + style)

Mark the axes: weights with "g" and lengths with "mm" unit of measure; 
remove those units from the axes' titles, replace "_"s with more 
readable " "s, capitalize the first letter of the axes' titles. Write 
them in blue using sans-serif (`sans`) font family.

```{r final-scale, exercise=TRUE, exercise.lines = 13}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = flipper_length_mm,
    label = species, colour = sex)) +
  geom_text(check_overlap = TRUE) +
  annotate("label", x = 4500, y = 170, label = "Short",
           hjust = "inward", vjust = "inward") +
  annotate("label", x = 4500, y = 230, label = "Long",
           vjust = "inward", hjust = "inward")

```


```{r final-scale-check}
grade_result(pass_if(~{
  ref <- penguins %>% 
    ggplot(aes(x = body_mass_g, y = flipper_length_mm,
      label = species, colour = sex)) +
    geom_text(check_overlap = TRUE) +
    annotate("label", x = 4500, y = 170, label = "Short",
             hjust = "inward", vjust = "inward") +
    annotate("label", x = 4500, y = 230, label = "Long",
             vjust = "inward", hjust = "inward") +
    scale_x_continuous("Body mass",
      labels = scales::label_comma(suffix = " g")) +
    scale_y_continuous("Flipper length",
      labels = scales::label_number(suffix = " mm")) +
    theme(axis.title = element_text(family = "sans", colour = "blue"))

  test_plots(.result, ref)
  
}), incorrect = "Wrong answer.")
```


```{r final-scale-hint-1, eval = FALSE}
# you can use
  scale_x_continuous(___, labels = scales::label_number(suffix = ___))
```

```{r final-scale-hint-2, eval = FALSE}
# you can use
  scale_x_continuous(___, labels = scales::label_number(suffix = ___))

# To style the axes' titles you can use the `them()` function
```

```{r final-scale-hint-3, eval = FALSE}
# you can use
  scale_x_continuous(___, labels = scales::label_number(suffix = ___))

# To style the axes' titles you can use the `theme()` function
theme(axis.___ = element_text(___ = "sans", colour = ___)
```



### Scale - legend (Meta-data + theme)

Capitalize the legend's title (replacing "_"s with " "s) and texts, and
made the legend's title dark green. 
   
```{r final-legend, exercise=TRUE, exercise.lines = 17}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = flipper_length_mm,
    label = species, colour = sex)) +
  geom_text(check_overlap = TRUE) +
  annotate("label", x = 4500, y = 170, label = "Short",
           hjust = "inward", vjust = "inward") +
  annotate("label", x = 4500, y = 230, label = "Long",
           vjust = "inward", hjust = "inward") +
  scale_x_continuous("Body mass",
    labels = scales::label_comma(suffix = " g")) +
  scale_y_continuous("Flipper length",
    labels = scales::label_comma(suffix = " mm")) +
  theme(axis.title = element_text(family = "sans", colour = "blue"))
```




```{r final-legend-check}
grade_result(pass_if(~{
  ref <- penguins %>% 
    ggplot(aes(x = body_mass_g, y = flipper_length_mm,
      label = species, colour = sex)) +
    geom_text(check_overlap = TRUE) +
    annotate("label", x = 4500, y = 170, label = "Short",
             hjust = "inward", vjust = "inward") +
    annotate("label", x = 4500, y = 230, label = "Long",
             vjust = "inward", hjust = "inward") +
    scale_x_continuous("Body mass",
      labels = scales::label_comma(suffix = " g")) +
    scale_y_continuous("Flipper length",
      labels = scales::label_comma(suffix = " mm")) +
    scale_colour_discrete("Sex",
      breaks = c("female", "male", NA),
      labels = c("Female", "Male", "NA")) +
    theme(axis.title = element_text(family = "sans", colour = "blue"),
          legend.title = element_text(colour = "darkgreen"))

  test_plots(.result, ref)
  
}), incorrect = "Wrong answer.")
```



```{r final-legend-hint-1, eval = FALSE}
# If you do not need to change colours, you should go with 
scale_colour_discrete() # and don't use the `value` option
```

```{r final-legend-hint-2, eval = FALSE}
# If you do not need to change colours, you should go with 
scale_colour_discrete() # and don't use the `value` option

scale_colour_discrete(
    breaks = c("female", "male", ___),
    labels = c(___, ___, ___))
```

```{r final-legend-hint-3, eval = FALSE}
# If you do not need to change colours, you should go with 
scale_colour_discrete() # and don't use the `value` option

# Remember to capitalize the labels and that NA _is_ a missing, while
# "NA" is a string
scale_colour_discrete(
    breaks = c("female", "male", NA),
    labels = c(___, ___, ___))
```

```{r final-legend-hint-4, eval = FALSE}
# If you do not need to change colours, you should go with 
scale_colour_discrete() # and don't use the `value` option

# Remember to capitalize the labels and that NA _is_ a missing, while
# "NA" is a string
scale_colour_discrete(
    breaks = c("female", "male", NA),
    labels = c(___, ___, "NA"))

# You can set the scale name (the legend's title in this case) both
# from the `scale_<AES>_<TYPE>()` function (first argument)
scale_colour_discrete(___,
    breaks = c("female", "male", NA),
    labels = c(___, ___, "NA"))
# OR you can set it using `labs()`, naming the corresponding aesthetics
labs(colour = ___) 
```

```{r final-legend-hint-5, eval = FALSE}
# If you do not need to change colours, you should go with 
scale_colour_discrete() # and don't use the `value` option

# Remember to capitalize the labels and that NA _is_ a missing, while
# "NA" is a string
scale_colour_discrete(
    breaks = c("female", "male", NA),
    labels = c(___, ___, "NA"))

# You can set the scale name (the legend's title in this case) both
# from the `scale_<AES>_<TYPE>()` function (first argument)
scale_colour_discrete(___,
    breaks = c("female", "male", NA),
    labels = c(___, ___, "NA"))
# OR you can set it using `labs()`, naming the corresponding aesthetics
labs(colour = ___) 

# To make the legend's title dark green you can set it "darkgreen"
theme(legend.___ = element_text(___ = ___))
```

   
   
   
   
   
### Facets (Theme + very little of meta-data)

Divide the plot into a grid of plots. Display the years into the columns 
and the islands into the rows. Keep the indication of variables in rows 
and columns' titles, and write them both in **bold** and _italic_.


```{r final-facet, exercise=TRUE}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = flipper_length_mm,
    label = species, colour = sex)) +
  geom_text(check_overlap = TRUE) +
  annotate("label", x = 4500, y = 170, label = "Short",
           hjust = "inward", vjust = "inward") +
  annotate("label", x = 4500, y = 230, label = "Long",
           vjust = "inward", hjust = "inward") +
  scale_x_continuous("Body mass",
    labels = scales::label_comma(suffix = " g")) +
  scale_y_continuous("Flipper length",
    labels = scales::label_comma(suffix = " mm")) +
  scale_colour_discrete("Sex",
    breaks = c("female", "male", NA),
    labels = c("Female", "Male", "NA")) +
  theme(axis.title = element_text(family = "sans", colour = "blue"),
        legend.title = element_text(colour = "darkgreen"))
```



```{r final-facet-check}
grade_result(pass_if(~{
  ref <- penguins %>% 
    ggplot(aes(x = body_mass_g, y = flipper_length_mm,
      label = species, colour = sex)) +
    geom_text(check_overlap = TRUE) +
    annotate("label", x = 4500, y = 170, label = "Short",
             hjust = "inward", vjust = "inward") +
    annotate("label", x = 4500, y = 230, label = "Long",
             vjust = "inward", hjust = "inward") +
    scale_x_continuous("Body mass",
      labels = scales::label_comma(suffix = " g")) +
    scale_y_continuous("Flipper length",
      labels = scales::label_comma(suffix = " mm")) +
    scale_colour_discrete("Sex",
      breaks = c("female", "male", NA),
      labels = c("Female", "Male", "NA")) +
    facet_grid(island~year, labeller = "label_both") +
    theme(axis.title = element_text(family = "sans", colour = "blue"),
          legend.title = element_text(colour = "darkgreen"),
          strip.text = element_text(face = "bold.italic"))

  test_plots(.result, ref)
  
}), incorrect = "Wrong answer.")
```


```{r final-facet-hint-1, eval = FALSE}
# Remember to keep the call to `theme()` as the last one
```

```{r final-facet-hint-2, eval = FALSE}
# Remember to keep the call to `theme()` as the last one

# For faceting, remember that `<LHS> ~ <RHS>` will display the <LHS> 
# variable across the grid rows (ie, every row is a distinct level
# of <LHS>), and `<RHS>` across the grid column
facet_grid(___~___)
```

```{r final-facet-hint-3, eval = FALSE}
# Remember to keep the call to `theme()` as the last one

# For faceting, remember that `<LHS> ~ <RHS>` will display the <LHS> 
# variable across the grid rows (ie, every row is a distinct level
# of <LHS>), and `<RHS>` across the grid column
facet_grid(___~___)

# You can use (the labeller is useful if you do not 
# want standard label, see `?labeller` for more information)
# to put the variable indication before the corresponding levels you can
# use `"label_both"`
facet_grid(___~___, labeller = ___)
```

```{r final-facet-hint-4, eval = FALSE}
# Remember to keep the call to `theme()` as the last one

# For faceting, remember that `<LHS> ~ <RHS>` will display the <LHS> 
# variable across the grid rows (ie, every row is a distinct level
# of <LHS>), and `<RHS>` across the grid column
facet_grid(___~___)

# You can use (the labeller is useful if you do not 
# want standard label, see `?labeller` for more information)
# to put the variable indication before the corresponding levels you can
# use `"label_both"`
facet_grid(___~___, labeller = ___)

# For bold and italic characters, you can use "bold.italic" where 
# needed in the theme
element_text(face = ___)
```

```{r final-facet-hint-5, eval = FALSE}
# Remember to keep the call to `theme()` as the last one

# For faceting, remember that `<LHS> ~ <RHS>` will display the <LHS> 
# variable across the grid rows (ie, every row is a distinct level
# of <LHS>), and `<RHS>` across the grid column
facet_grid(___~___)

# You can use (the labeller is useful if you do not 
# want standard label, see `?labeller` for more information)
# to put the variable indication before the corresponding levels you can
# use `"label_both"`
facet_grid(___~___, labeller = ___)

# For bold and italic characters, you can use "bold.italic" where 
# needed in the theme
theme(strip.text = element_text(face = "bold.italic"))
```

   
   



### Theme (stytle)

Change the plot title to "Penguins' mass and their flipper length by
years", add a caption to acknowledge that "Data from {palmerpenguins}
package.". Finally, enlarge the title to 24 points and make it orange.




```{r final-theme, exercise=TRUE}
penguins %>% 
  ggplot(aes(x = body_mass_g, y = flipper_length_mm,
    label = species, colour = sex)) +
  geom_text(check_overlap = TRUE) +
  annotate("label", x = 4500, y = 170, label = "Short",
           hjust = "inward", vjust = "inward") +
  annotate("label", x = 4500, y = 230, label = "Long",
           vjust = "inward", hjust = "inward") +
  scale_x_continuous("Body mass",
    labels = scales::label_comma(suffix = " g")) +
  scale_y_continuous("Flipper length",
    labels = scales::label_comma(suffix = " mm")) +
  scale_colour_discrete("Sex",
    breaks = c("female", "male", NA),
    labels = c("Female", "Male", "NA")) +
  facet_grid(island~year, labeller = "label_both") +
  theme(axis.title = element_text(family = "sans", colour = "blue"),
        legend.title = element_text(colour = "darkgreen"),
        strip.text = element_text(face = "bold.italic"))
```




```{r final-theme-check}
grade_result(pass_if(~{
  ref <- penguins %>% 
    ggplot(aes(x = body_mass_g, y = flipper_length_mm,
      label = species, colour = sex)) +
    geom_text(check_overlap = TRUE) +
    annotate("label", x = 4500, y = 170, label = "Short",
             hjust = "inward", vjust = "inward") +
    annotate("label", x = 4500, y = 230, label = "Long",
             vjust = "inward", hjust = "inward") +
    scale_x_continuous("Body mass",
      labels = scales::label_comma(suffix = " g")) +
    scale_y_continuous("Flipper length",
      labels = scales::label_comma(suffix = " mm")) +
    scale_colour_discrete("Sex",
      breaks = c("female", "male", NA),
      labels = c("Female", "Male", "NA")) +
    facet_grid(island~year, labeller = "label_both") +
    labs(title = "Penguins' mass and their flipper length by years",
         caption = "Data from {palmerpenguins} package.") +
    theme(plot.title = element_text(size = 24, colour = "orange"),
          axis.title = element_text(family = "sans", colour = "blue"),
          legend.title = element_text(colour = "darkgreen"),
          strip.text = element_text(face = "bold.italic"))

  test_plots(.result, ref)
  
}), incorrect = "Wrong answer.")
```



```{r final-theme-hint-1, eval = FALSE}
# You can add main titles (plot title, caption and all the title's
# legend for the aesthetics) using `labs()`
labs(title = ___,  caption = ___)
```

```{r final-theme-hint-2, eval = FALSE}
# You can add main titles (plot title, caption and all the title's
# legend for the aesthetics) using `labs()`
labs(title = ___,  caption = ___)

# You can adjust the theme of the plot title as usual
theme(___.title = element_text(___ = 24, ___ = "orange"))
```



### Greetings!!!

We agree our final plot is not lovely, nor insightful... but our excuse 
for that is: it was only a _pre**text**_ plot ;-)

###

You are now a master of text and labels into `{ggplot2}`.
Remember to distinguish when your "text" requirements come from:

- **data**: map the corresponding variable to the `label` aesthetic,
  and draw them with `geom_text()` or `geom_label()` at your convenience

- **meta-data**: maybe that means you want to manage some title or text
  in some legends (or axis, which are nothing more than the axes' 
  legend). So, you can go with the corresponding scale_<AES>_<TYPE>()`
  function (which the first argument is the legend title), and you can 
  use the `break`, and `label` to control which label to plot and their
  corresponding name respectively. Or, just to change the title, you can
  use the convenient wrapper `labs()` just assigning the title you like 
  to the corresponding aesthetics (eg, labs(colour = "Colour legend")).

- **just an annotation**: so, you simply want to draw a `geom_*` for a 
  single-new not-in-data point ;-)... use the `annotation()` function,  
  and in its first argument, set the geom you like (eg, "text" or 
  "label"), next directly assign the value you need (eg, `x`, `y`,  
  `label`).

- **style**: maybe you already have some text you like to make more 
  stylish. So, you need to put a call to `theme()` and adjust the 
  corresponding elements' characteristics. For text elements they are
  set calling `element_text()` for the corresponding argument
  of interest (eg, `plot.title`).
  
  
That's all!!
Congratulations!!
  
Thank you for having been part of this class.
  
___\@CorradoLanera___


Feel free to play with penguins and `{ggplot2}` as much as you like!

```{r final-free, exercise=TRUE, exercise.lines = 5}
library(tidyverse)
library(palmerpenguins)



```







```{r quiz-template, eval=FALSE, include=FALSE}
quiz(caption = "XXX-quiz",
  question("q",
           
    answer("a",
           correct = TRUE,
           message = random_praise()),
    
    answer("b",
           message = ""),
    
    answer("c",
           message = ""),
    
    answer("d",
           message = ""),
    
    try_again = random_encouragement(),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```


